Intersection of non-matching grids of different dimensions

### 2016-09-29 paper_intersections.tex

We consider a composed mesh $\mathcal T$ consisting of simplicial meshes $\mathcal T_i$ of dimensions $d_i \in \{1,2,3\}$, $i=1,\dots,N_\mathcal{T}$
in the 3d ambient space. We assume that every mesh $\mathcal T_i$ is a connected set with no self intersection.
Further we assume only single 3d mesh $\mathcal{T}_1$.
The mesh intersection problem is to find all pairs of elements $L\in \mathcal{T}_i$, $K\in \mathcal{T}_j$, $i\ne j$ that have non-empty intersection 
and compute that intersection.

The mesh intersection problem consists of the two parts: First, generate a set of candidate pairs $(K, L)$. Second, compute the intersection for particular pair.
In order to get candidate pairs efficiently, the existing algorithms use either various space trees \cite{massing_efficient_2013} or front tracing
\cite{gander_algorithm_2013} or both \cite{elsheikh_reliable_2012} as in our approach. To compute the actual intersections, we 
use the \plucker coordinates for the line-triangle intersections and a modification of the  Platis and Theoharis algorithm \cite{platis_fast_2003}
for the line-tetrahedron intersections. These are used as the building block for the triangle-triangle and the triangle-tetrahedron cases.

@@

Our contribution is twofold:  First, we use the front tracking algorithm both to minimize set of intersection candidates as well 
to reuse part of calculations made on neighbouring intersections. Second, we present family of efficient algorithms based on \plucker coordinates 
for computing 1d-2d, 1d-3d, 2d-2d, and 2d-3d intersections of simplicies. The paper is organized as follows \dots

@@

\subsection{Prolongation algorithm}
Consider now a complex mesh of combined dimensions consisting of \emph{components}, which are sets of connected
elements of the same lower dimension (1D, 2D), in the space of 3D elements, which we shall call a \emph{bulk}. 
Obtaining all of the 1D-3D and 2D-3D intersections is based on finding the first two elements intersecting each other.
Then we can prolongate the intersection by investigating neighbouring elements.
%
\begin{figure}[!htb]
%   \vspace{0pt}
  \centering    
    \includegraphics[width=\textwidth]{\figpath prolongation_scheme.pdf}
  \caption{Prolongation algorithm for 1D-2D and 2D-3D intersections. }
  \label{fig:prolongation}
\end{figure}

The prolongation algorithm is the same for both 1D and 2D components, see \fig{fig:prolongation}. 
It can be seen as a \emph{breadth-first search}
\footnote{Wiki, [online 2016-03-01], \url{https://en.wikipedia.org/wiki/Breadth-first_search}}
algorithm on the graph of component elements.

\textbf{breadth first search} algorithm:
  \begin{enumerate}
    \item\label{en:first} Get next unprocessed component element $k$.
    \item Find intersection candidates $\mathcal K$ in bulk mesh (3D elements).
    \item \label{en:q2}For $K\in \mathcal K$ compute intersection $(k, K)$.  
    \item Push the intersection neigbours into queues: $(k, L) \to Q_b$, $(l, K) \to Q_c$. 
    \item While $(l, L) \in Q_b$ check intersection $(l, L)$. Append queues.
    \item Pop pair from $Q_c$. (move to the next component element: go to \ref{en:q2}).
    \item If $Q_c$ is empty, go to \ref{en:first}.
  \end{enumerate}
  Prerequisites:
  \begin{itemize}
   \item need information about element connectivity (mesh preprocessing)
%   \item mark and check already inspected pairs. Marking elements + hashing.
   \item find the starting intersection -- AABB (Axes Aligned Bounding Boxes) 
   \item possibly create Bounding Interval Hierarchy (BIH)
%   \item Pass already known information with pairs.
%   \item Deal with degenerated cases.
  \end{itemize}

After computing the intersection of a pair of elements (line or triangle vs tetrahedron), we fill
two queues with element pairs as candidates for further intersection. If the intersection edge 
(point of line in 1D, edge of polygon in 2D) is inside the tetrahedron, not on its surface, we
get a neighbouring element of the component and push it back together with the current tetrahedron into 
\emph{component prolongation queue}. If the intersection edge is inside the \emph{slave} element 
(line or triangle), i.e. is on the surface of tetrahedron, we get a neighbouring element of the tetrahedron
and push it back together with the current slave element into \emph{3D prolongation queue}.

Then we empty the prolongation queues -- the 3D prolongation queue at first, then the component prolongation queue.
When both queues are empty, all intersections of a component have been found and we continue to look for
another component.

The algorithm is now unified for 1D and 2D in contrast to \cite{fris_dp_2015}, where the component prolongation
queue is emptied at first.

@@

A new candidate pair of elements is found during prologantion, based on the topological information of the intersection corner. There are 3 possible cases (applies both for 1D-3D and 2D-3D):
\begin{itemize}
    \item \textbf{IC lies on component element side and inside bulk element} \\
            We find all the sides of component element in which the IC lies (IC can be at node and connect more sides).
            We find the component neighboring elements over the sides and push a new candidate pair
            [component neighbor -- current bulk element] into component queue.
    \item \textbf{IC lies on component element side and on the surface of bulk element} \\
            We find all the faces of bulk element in which the IC lies (1 face, or 2 faces (IC on an edge), 
            or 3 faces (IC at a node)). We find the corresponding neighboring bulk elements over the faces and
            push the following new candidates pairs into the bulk queue: [current component element -- bulk neighbor], [component neighbor -- bulk neighbor].
            If the candidates pair has been investigated already, we skip it.
    \item \textbf{IC lies inside component element (therefore must be on the surface of bulk element)} \\
            The same as before, but we push only [current component element -- bulk neighbor] candidates pairs, since
            there is no component neighbor.
\end{itemize}

>>>>

We consider a composed mesh $\mathcal T$ consisting of simplicial meshes $\mathcal T_i$ of dimensions $d_i \in \{1,2,3\}$, $i=1,\dots,N_\mathcal{T}$
in the 3d ambient space. We assume that every mesh $\mathcal T_i$ is a connected set with no self intersection.
Further we assume only single 3d mesh $\mathcal{T}_1$.
The mesh intersection problem is to find all pairs of elements $L\in \mathcal{T}_i$, $K\in \mathcal{T}_j$, $i\ne j$ that have non-empty intersection 
and to compute that intersection.

The mesh intersection problem consists of the two parts: First, generate a set of candidate pairs $(K, L)$. Second, compute the intersection for particular pair.
In order to get candidate pairs efficiently, the existing algorithms use either various space trees \cite{massing_efficient_2013} or front tracing
\cite{gander_algorithm_2013} or both \cite{elsheikh_reliable_2012} as in our approach. To compute the actual intersections, we 
use the \plucker coordinates for the line-triangle intersections and a modification of the  Platis and Theoharis algorithm \cite{platis_fast_2003}
for the line-tetrahedron intersections. These are used as the building blocks for the triangle-triangle and the triangle-tetrahedron cases.

@@

\notePE{Unify the term 'front tracing' 'front tracking' 'advancing front'.}
Our contribution is twofold:  First, we use the front tracking algorithm both to minimize set of intersection candidates as well 
to reuse part of calculations made on neighbouring intersections. Second, we present family of efficient algorithms based on \plucker coordinates 
for computing 1d-2d, 1d-3d, 2d-2d, and 2d-3d intersections of simplicies. The paper is organized as follows \dots

@@

\section{Front tracking algorithm}
Prerequisites:
  \begin{itemize}
   \item need information about element connectivity (mesh preprocessing)
%   \item mark and check already inspected pairs. Marking elements + hashing.
   \item find the starting intersection -- AABB (Axes Aligned Bounding Boxes) 
   \item possibly create Bounding Interval Hierarchy (BIH)
%   \item Pass already known information with pairs.
%   \item Deal with degenerated cases.
  \end{itemize}

Consider now a complex mesh of combined dimensions consisting of \emph{components}, which are sets of connected
elements of the same lower dimension (1D, 2D), in the space of 3D elements, which we shall call a \emph{bulk}. 
Obtaining all of the 1D-3D and 2D-3D intersections is based on finding the first two elements intersecting each other. Then we can prolongate the intersection by investigating neighbouring elements.

It can be seen as a \emph{breadth-first search}
\footnote{Wiki, [online 2016-03-01], \url{https://en.wikipedia.org/wiki/Breadth-first_search}}
algorithm on the graph of component elements.

\textbf{breadth first search} algorithm:
  \begin{enumerate}
    \item\label{en:first} Get next unprocessed component element $k$.
    \item Find intersection candidates $\mathcal K$ in bulk mesh (3D elements).
    \item \label{en:q2}For $K\in \mathcal K$ compute intersection $(k, K)$.  
    \item Push the intersection neigbours into queues: $(k, L) \to Q_b$, $(l, K) \to Q_c$. 
    \item While $(l, L) \in Q_b$ check intersection $(l, L)$. Append queues.
    \item Pop pair from $Q_c$. (move to the next component element: go to \ref{en:q2}).
    \item If $Q_c$ is empty, go to \ref{en:first}.
  \end{enumerate}

%
\begin{figure}[!htb]
%   \vspace{0pt}
  \centering    
    \includegraphics[width=\textwidth]{\figpath prolongation_scheme_ip.pdf}
  \caption{Prolongation algorithm for 1D-2D and 2D-3D intersections. }
  \label{fig:prolongation}
\end{figure}

The advancing front algorithm works the same way for both 1D and 2D components and is displayed in \fig{fig:prolongation}. The main idea is to compute intersections for a component element with all 
possible bulk elements, and then move to a~next neighboring component element. For this reason, we define two
queues of candidate pairs: a~\emph{bulk queue} and a~\emph{component queue} (yellow boxes in \fig{fig:prolongation}). We shall now describe the algorithm in detail.

On input we consider a candidate pair, for which a non-empty intersection is computed.
Now we look for new candidate pairs among the neighboring elements (the block Prolongation decide).
Therefore, we iterate over the intersection points and further exploit their topological information.
There are 3 possible cases (applies both for 1D-3D and 2D-3D):
\begin{itemize}
    \item \textbf{IP lies on the component element side and inside the bulk element} \\
            We find all the sides of component element in which the IP lies (IP can be at node and connect more sides).
            Next, we find the component neighboring elements over the sides and push all new candidate pairs
            [component neighbor -- current bulk element] into the component queue.
            Note, that there can be more than one neighbor on a side, if the component has branches.
    \item \textbf{IP lies on the component element side and on the surface of the bulk element} \\
            We find all the faces of bulk element in which the IP lies (1 face, or 2 faces (IP on an edge), 
            or 3 faces (IP at a node)). We find the corresponding neighboring bulk elements over the faces and
            push the following new candidate pairs into the bulk queue: [current component element -- bulk neighbor], [component neighbor -- bulk neighbor].
    \item \textbf{IP lies inside component element (therefore must be on the surface of bulk element)} \\
            We proceed as in previous case, but we push only [current component element -- bulk neighbor] candidate pairs, since there is no component neighbor.
\end{itemize}

If the candidate pair has been found already, we skip it.
We also see that the candidate pairs are of three types:
[current component element -- bulk neighbor],
[component neighbor -- current bulk element], 
[component neighbor -- bulk neighbor],
from which only the first one goes into the bulk queue, trying to cover the whole component element.

% After computing the intersection of a pair of elements (line or triangle vs tetrahedron), we fill
% two queues with element pairs as candidates for further intersection. If the intersection edge 
% (point of line in 1D, edge of polygon in 2D) is inside the tetrahedron, not on its surface, we
% get a neighbouring element of the component and push it back together with the current tetrahedron into 
% \emph{component prolongation queue}. If the intersection edge is inside the \emph{slave} element 
% (line or triangle), i.e. is on the surface of tetrahedron, we get a neighbouring element of the tetrahedron
% and push it back together with the current slave element into \emph{3D prolongation queue}.

Then we empty the two queues. 
We pop out new candidate pairs from the \emph{bulk queue} as long as it is not empty and for every new intersection computed, we repeat the previous part (means that we can further fill both queues).
The \emph{bulk queue} is empty when the component element is fully covered by bulk elements, or
when there is no bulk neighbor to which we can advance.
Then we can pop a new candidate pair from \emph{component prolongation queue} and process it.
When both queues are empty, all intersections of a component have been found and we start over by looking for the first intersection of another component.

% The algorithm is now unified for 1D and 2D in contrast to \cite{fris_dp_2015}, where the component prolongation
% queue is emptied at first.

@@
